---
title: "Example 1: House price forecast"
author: "Author: Gurakuqi Jurgen"
output:
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```  


The dataset houses-canada.csv includes 4 variables about a random selection of 546 houses sold in a Canadian city:
price: Sale price of the house;

lotsize: Lot size of the property in square feet;

stories: Number of stories (floors) excluding basement;

gasheat: Does the house use gas for hot water heating?


First, we read the data, then we show 6 rows of the dataset to understand how
the variables of the dataset are populated:
```{r }
house <- read.csv("houses-canada.csv")
head(house)
```



We also print the data to understand how the data is distributed and if some 
variables need to be factorized:
```{r }
summary(house)
table(house$stories)
```  



As we can see, there are categorical variables which need to be transformed into factors:
```{r }
house$gasheat <- factor(house$gasheat)
house$stories <- factor(house$stories)
summary(house)
```


We try now to fit a possible model:
```{r }
mod1 <- lm(price ~ lotsize + stories + gasheat, data = house)
summary(mod1)
```
As we can see in the summary, all the predictors, except gasheat are strongly 
significant and positively related to the response. Gasheat is also positively
related to price, but just significant.
Also the intercept, which includes stories1, is strongly significant and 
positively related to the response.
The $R^2$ is *44%*, and the adjusted $R^2$ is *43%*.


Now, we must validate the model assumptions through the residual analysis and 
the normal probability plot analysis:
```{r }
library(car)
residualPlots(mod1)
qqPlot(residuals(mod1))
```
We can see that:

- The qqplot shows a strong departure from the normality on the right.
- The residuals show a lot of outliers for gasheat and stories, and some patterns
on the lotsize and on the fitted values. The problems are also confirmed by the 
weakly significant Tukey test, and by the strongly significant test statistic of
lotsize, so something is missing.
  

We can also check the effect plots to check the relation between the predictors
and the response, and confirm what we have seen in the summary:
```{r }
library(effects)
plot(allEffects(mod1))
```
Every effect plot confirms what we have previously said about the predictors in
the summary:

- Lotsize causes strong increments in the price according to its increase.
- Stories causes great changes in the price according to it's value change.
- Gasheat causes a visible increase when changing from no to yes.


We check for the presence of possible *influential points*:
```{r }
influenceIndexPlot(mod1, vars = "Cook")
```

The Cook's distance tell's us that there are not influential points, as none of 
the points gets below threshold of 0.5.

Also we check for the presence of multicollinearities:
```{r }
vif(mod1)
```
The possible multicollinearities are indicated for values near to or higher than
5, and strong multicollinearities are inidicated for values higher than 10, but
in this case none of the values shows problems, so there are not multicollinarities.

A last check we do is about the values that this model can return, considering
that a money response should not assume negative values:
```{r }
summary(fitted(mod1))
```
This model does not show problems about negative responses, so we do not need
to fit a logarithmic version of it.


We could try to improve our model with the introduction of an interaction term:
```{r }
mod2 <- update(mod1, . ~ . + stories * lotsize)
summary(mod2)
```
As we can see in the summary:

- **lotsize** is *strongly* significant, and *positively* related to the response.
- **stories** has become *not* significant, and it's *positively* related to the
response.
- **gasheat** has mantained it's significance and positive relation with price.
- The interaction **lotsize/stories** is *weakly* significant for stories 2, and not significant
for 3 and 4, whereas the 1 is included in the intercept, which has remained
strongly significant.
- Stories alone seems to be not significant, but yet positively related to price.
- The intercept, which includes stories1 and stories1:lotsize, is strongly
significant and positively related to the response.
- The $R^2$ is *45%*, and the adjusted $R^2$ is *44%*, which 
are better then those of the previous model, but not that much.



Again, we check the residuals and the normal probability plot:
```{r }
residualPlots(mod2)
qqPlot(residuals(mod2))
```
Yet, the qqplot shows a strong departure from the normality, and the residuals
are not good again. Also, this time the Tukey test has got worse then before.
  
We check again the effect plots:
```{r }
plot(allEffects(mod2))
```
Each plot confirms the previous results shown in the summary:

- Gasheat is significant, as it rises the price by changing from No to yes, but
not in a concrete way.
- The inreaction does not show clear increases for stories 3 and 4, but shows
some light increases for 1 and 2.


We can try also a last model using other interactions:
```{r }
mod3 <- update(mod2, . ~ . + gasheat * lotsize)
summary(mod3)
```
As we can see in the summary:

- **lotsize** is *strongly* significant, and *positively* related to the response.
- **stories** is *not* significant, and it's *positively* related to the
response.
- **gasheat** has become not significant, and it's positively related with price.
- The interaction **lotsize/stories** is *weakly* significant for stories 2, and not significant
for 3 and 4, whereas the 1 is included in the intercept, which has remained
strongly significant.
- The interaction **lotsize/gasheat** is *not* significant 2, and it's psotively
related to the response.
- Stories is not significant, but yet positively related to price.
- The intercept, which includes stories1 and stories1:lotsize and
lotsize:gasheatNo, is strongly significant and positively related to the response.
- The $R^2$ is *45%*, and the adjusted $R^2$ is *44%*, which 
are better then those of the previous model, but not that much.



Again, we check the residuals and the normal probability plot:
```{r }
residualPlots(mod3)
qqPlot(residuals(mod3))
```
The qqplot shows again a strong departure from the normality, and the residuals
are not good.
Also here the test significance for lotsize and the Tukey test are both strongly
significant.
  
We check again the effect plots:
```{r }
plot(allEffects(mod3))
```
The interaction between lotsize and stories behaves like in the previous model,
whereas the interaction between lotsize and gasheat shows that for Gasheat
"yes" there is no clearly strong increment of price, but there is a visible
increment when gasheat is "no".


Eventually, we pick the *mod1* as final model for our predictions, because 
even if it shows a lightly lower $R^2$ and adjuted $R^2$, it shows less problems
with lotsize, has not significant interaction terms, has less variables 
(meaning a better option according to the Parsimony principle).

We compute the predictions and predictions intervals for mod1:
```{r }
pr <- predict(mod1, 
              newdata = data.frame(gasheat = "yes", 
                                   stories = "1",
                                   lotsize = 3000), 
              se.fit = TRUE)
ci <- pr$fit + c(-1, 1) * qnorm(0.975) * pr$se
pr$fit
ci

pr <- predict(mod1, 
              newdata = data.frame(gasheat = "no", 
                                   stories = "2",
                                   lotsize = 3000),  
              se.fit = TRUE)
ci <- pr$fit + c(-1, 1) * qnorm(0.975) * pr$se
pr$fit
ci

pr <- predict(mod1, 
              newdata = data.frame(gasheat = "yes", 
                                   stories = "4",
                                   lotsize = 3000), 
              se.fit = TRUE)
ci <- pr$fit + c(-1, 1) * qnorm(0.975) * pr$se
pr$fit
ci

pr <- predict(mod1, 
              newdata = data.frame(gasheat = "yes", 
                                   stories = "1",
                                   lotsize = 6000), 
              se.fit = TRUE)
ci <- pr$fit + c(-1, 1) * qnorm(0.975) * pr$se
pr$fit
ci
```



 